Neighbor-Based Decentralized Training Strategies for Multi-Agent Reinforcement Learning

### 2024-10-02 paper-2024-sac-marl-neighbouring-strategies.tex
chore: continue in refine main concepts

\section{Background and Motivation}\label{sec:background}
Multi-agent reinforcement learning is an extension of a sequientail decision-making process, 
where multiple agents interact with each other and the environment.
In particular, each agent in the system aims to maximize its own reward,
which is typically a function of the global state (sometime not known, aka partial observable) and the joint action taken by all agents.
%
In doing that, each agent perceive the environment through its own observation function,
which maps the global state to the agent's local observation and perform an action based on its policy.
%
This action is then combined with the actions of other agents to form a joint action,
which is used to update the global state and reward.

MADRL is a subfield of reinforcement learning that focuses on training multiple agents to interact with each other within a shared environment.
%
In this \emph{deep} variant, agents are typically represented by neural networks that learn to maximize a shared (or individual) reward signal.
% 
%
MADRL is a wide area, and it encompassed a variety of scenarios, 
such as cooperative (e.g., multi-agent pathfinding), 
competitive (e.g., multi-agent games), 
and mixed environments.
%
In this paper, we focus on cooperative scenarios, where agents must collaborate to achieve a common goal,
introducing a common formalization (networked agents) and discussing the key concepts of joint action space, shared reward, partial observability, and credit assignment, following by different training schemes (centralized and distributed), with a discussion on the state of the art and motivation for our work.
\subsection{Formalization}
Since MARL is a broad field, it is essential to provide a formalization to guide the discussion.
In this paper, we consider a partially observable networked multi-agent system (PONMAS) as a tuple $(\mathcal{G}, \mathcal{S}, \mathcal{A}, \mathcal{O}, \mathcal{P}, \mathcal{R}, \gamma)$, where:
$\mathcal{G} = (N, E)$ is a communication graph, where $N = {1, \dots, n}$ is the set of $n$ agents and $E \subseteq N \times N$ represents the communication links between agents. Time-varying graphs $\mathcal{G}_t = (N, E_t)$ can be used to represent communication evolving over time $t$.
%
$\mathcal{S}$ is the global state space.
%
$\mathcal{A} = \mathcal{A}^1 \times \dots \times \mathcal{A}^n$ is the joint action space, where $\mathcal{A}^i$ is the action space of agent $i$.
%
$\mathcal{O} = \mathcal{O}^1 \times \dots \times \mathcal{O}^n$ is the joint observation space, where $\mathcal{O}^i$ is the observation space for agent $i$. 
%
Observations $o^i \in \mathcal{O}^i$ provide partial information about the global state.
$\mathcal{P}: \mathcal{S} \times \mathcal{A} \times \mathcal{S} \to [0, 1]$ is the state transition function, describing the probability of transitioning to a new state $s' \in \mathcal{S}$ given the current state $s \in \mathcal{S}$ and joint action $a \in \mathcal{A}$.
%
$\mathcal{R} = {\mathcal{R}^i}_{i \in N}$, where $\mathcal{R}^i: \mathcal{S} \times \mathcal{A} \to \mathbb{R}$ is the reward function for agent $i$.
%
$\gamma \in [0, 1]$ is the discount factor.
%
Each agent $i$ at time $t$ receives an observation $o^i_t \in \mathcal{O}^i$, takes an action $a^i_t \in \mathcal{A}^i$ based on its local policy $\pi^i: \mathcal{O}^i \times \mathcal{A}^i \to [0,1]$, and receives a reward $r^i_{t+1} = \mathcal{R}^i(s_t, a_t)$. 
%
The global state evolves according to $\mathcal{P}$. 
Agents can exchange information with their neighbors in $\mathcal{G}$ (or $\mathcal{G}t$ if the graph is time-varying). 
%
The objective is typically to maximize the expected discounted sum of rewards $\sum{t=0}^{\infty} \gamma^t r^i_t$ for each agent $i$, 
or some global reward function like the average reward over all agents.
Formally, it can be defined as:
\[ J(\{\pi^i\}) = \lim_{T \to \infty} \mathbb{E} \left[ \frac{1}{T} \sum_{t=0}^{T-1} \frac{1}{N} \sum_{i \in \mathcal{N}} R^i(s_t, a_t) \right] \]

@@

Similar to the individual reward function, a global average reward function can be defined as $\bar{\mathcal{R}}(s, a) = \frac{1}{n} \sum_{i=1}^n \mathcal{R}^i(s,a)$, leading to a global average Q-value:

\[ \bar{Q}(s, a) = \mathbb{E}_{\pi} \left[ \sum_{t=0}^{\infty} \gamma^t \bar{\mathcal{R}}(s_t, a_t) | s_0 = s, a_0 = a \right] \]

where $\pi = \{\pi^i\}_{i \in N}$ represents the joint policy. 
The goal in collaborative MARL is often to find a joint policy that maximizes this global average Q-value or a similar global objective.
There are several algorithms which try to find the optimal Q-value function, one of the most popular is the deep q-learning algorithm.
\subsubsection{Deep Q-Learning Formalization}
deep q-learning utilizes a deep neural network, 
often referred to as a Q-network, to approximate the action-value function $Q(s, a)$. 
This network takes the state $s$ as input and outputs a vector of Q-values, one for each possible action $a \in \mathcal{A}$.  
The optimal action-value function, denoted $Q^*(s, a)$, satisfies the Bellman optimality equation:

@@

where $\theta$ are the parameters of the Q-network, $\theta^-$ are the parameters of a target Q-network (used for stability), 
and $\mathcal{D}$ is the experience replay buffer, a collection of past experiences $(s, a, r, s')$.
%
In the multi-agent setting, each agent $i$ has its own Q-network $Q^i(o^i, a^i; \theta^i)$ that takes its local observation $o^i$ as input.  
For centralized training, the global state $s$ can be used as input. 
The specific form of the multi-agent Q-function and the update rules depend on the chosen MARL algorithm (e.g., independent Q-learning, value decomposition methods).  
For example, in independent deep Q-learning, each agent updates its Q-network independently using its local reward $r^i$:
\begin{equation}
  \resizebox{\linewidth}{!}{$
  \mathcal{L}^i(\theta^i) = \mathbb{E}_{(o^i, a^i, r^i, o'^i) \sim \mathcal{D}^i} \left[ \left( r^i + \gamma \max_{a'^i} Q^i(o'^i, a'^i; \theta^{i-}) - Q^i(o^i, a^i; \theta^i) \right)^2 \right]
  $}
  \end{equation}
%\subsection{Key Concepts}

>>>>

\section{Background and Motivation}\label{sec:background}
MARL is an extension of a sequential decision-making process, 
where multiple \emph{agents} interact with each other and a shared \emph{environment}.
In particular, each agent in the system aims to maximize its own \emph{reward signal} (i.e., a scalar feedback signal),
which is typically a function of the global state (sometime not known, making the system \emph{partial observable}) and the joint action taken by all agents.
%
In doing that, each agent perceive the environment through its own \emph{observation function},
which maps the global state to the agent's local observation and perform an action based on its \emph{policy} (i.e., a mapping from observations to actions).
%
This chosen action is then combined with the actions of other agents forming a \emph{joint action},
which is used to update the global state and reward.

In this paper, we focus on cooperative MADRL scenarios (so the deep version of MARL), 
where agents must collaborate to achieve a common goal.
In the following, we provide an overview of the key concepts in MADRL, 
including the formalization of the problem, the deep Q-learning algorithm, and the training paradigms used in the literature.
\subsection{Formalization}
Since MADRL is a broad field, it is essential to provide a formalization to guide the discussion.
In this paper, we consider partially observable networked markov decision process (PONMDP)~\cite{DBLP:journals/tac/AdlakhaLG12} as a tuple $(\mathcal{G}, \mathcal{S}, \mathcal{A}, \mathcal{O}, \mathcal{P}, \mathcal{R}, \gamma)$, where

\begin{itemize}
  \item $\mathcal{G} = (N, E)$ is a communication graph, where $N$ is the set of $n$ agents and $E \subseteq N \times N$ represents the communication links between agents. Time-varying graphs $\mathcal{G}_t = (N, E_t)$ can be used to represent communication evolving over time $t$.
  \item $\mathcal{S}$ is the global state space.
  \item $\mathcal{A} = \mathcal{A}^1 \times \dots \times \mathcal{A}^n$ is the joint action space, where $\mathcal{A}^i$ is the action space of agent $i$.
  \item $\mathcal{O} = \mathcal{O}^1 \times \dots \times \mathcal{O}^n$ is the joint observation space, where $\mathcal{O}^i$ is the observation space for agent $i$.
  \item $\mathcal{P}: \mathcal{S} \times \mathcal{A} \times \mathcal{S} \to [0, 1]$ is the state transition function, describing the probability of transitioning to a new state $s' \in \mathcal{S}$ given the current state $s \in \mathcal{S}$ and joint action $a \in \mathcal{A}$.
  \item $\mathcal{R} = {\mathcal{R}^i}_{i \in N}$, where $\mathcal{R}^i: \mathcal{S} \times \mathcal{A} \to \mathbb{R}$ is the reward function for agent $i$.
  \item $\gamma \in [0, 1]$ is the discount factor.
\end{itemize}
Each agent $i$ at time $t$ receives an observation $o^i_t \in \mathcal{O}^i$, 
takes an action $a^i_t \in \mathcal{A}^i$ based on its local policy $\pi^i: \mathcal{O}^i \times \mathcal{A}^i \to [0,1]$, and receives a reward $r^i_{t+1} = \mathcal{R}^i(s_t, a_t)$. 
%
The global state evolves according to $\mathcal{P}$. 
Agents can exchange information with their neighbors in $\mathcal{G}$ 
(or $\mathcal{G}t$ if the graph is time-varying). 
%
The objective is typically to maximize the expected discounted sum of rewards $\sum{t=0}^{\infty} \gamma^t r^i_t$ for each agent $i$, 
or some global reward function like the average reward over all agents.
Formally, it can be defined as:
\[ J(\{\pi^i\}) = \lim_{T \to \infty} \mathbb{E} \left[ \frac{1}{T} \sum_{t=0}^{T-1} \frac{1}{N} \sum_{i \in \mathcal{N}} R^i(s_t, a_t) \right] \]

@@

%Similar to the individual reward function, a global average reward function can be defined as $\bar{\mathcal{R}}(s, a) = \frac{1}{n} \sum_{i=1}^n \mathcal{R}^i(s,a)$, leading to a global average Q-value:

%\[ \bar{Q}(s, a) = \mathbb{E}_{\pi} \left[ \sum_{t=0}^{\infty} \gamma^t \bar{\mathcal{R}}(s_t, a_t) | s_0 = s, a_0 = a \right] \]

%where $\pi = \{\pi^i\}_{i \in N}$ represents the joint policy. 
%The goal in collaborative MARL is often to find a joint policy that maximizes this global average Q-value or a similar global objective.
There are several algorithms which try to find the optimal Q-value function, one of the most popular is the deep Q-learning algorithm.
\subsubsection{Deep Q-Learning Formalization}
deep Q-learning utilizes a deep neural network, 
often referred to as a Q-network, to approximate the action-value function $Q(s, a)$. 
This network takes the state $s$ as input and outputs a vector of Q-values, one for each possible action $a \in \mathcal{A}$.  
The optimal action-value function, denoted $Q^*(s, a)$, satisfies the Bellman optimality equation:

@@

where $\theta$ are the parameters of the Q-network, $\theta^-$ are the parameters of a target Q-network (used for stability), 
and $\mathcal{D}$ is the experience replay buffer, a collection of past experiences $(s, a, r, s')$.
%
In the multi-agent setting, each agent $i$ has its own Q-network $Q^i(o^i, a^i; \theta^i)$ that takes its local observation $o^i$ as input.  
For centralized training, the global state $s$ can be used as input. 
The specific form of the multi-agent Q-function and the update rules depend on the chosen MARL algorithm (e.g., independent Q-learning, value decomposition methods~\cite{DBLP:conf/icml/SonKKHY19}).  
For example, in independent deep Q-learning, each agent updates its Q-network independently using its local reward $r^i$:
\begin{equation}
  \resizebox{\linewidth}{!}{$
  \mathcal{L}^i(\theta^i) = \mathbb{E}_{(o^i, a^i, r^i, o'^i) \sim \mathcal{D}^i} \left[ \left( r^i + \gamma \max_{a'^i} Q^i(o'^i, a'^i; \theta^{i-}) - Q^i(o^i, a^i; \theta^i) \right)^2 \right]
  $}
  \end{equation}
%\subsection{Key Concepts}
